%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}

\setlength{\parskip}{5pt}%
%\setlength{\parindent}{0pt}%


%In addition, it would not be possible to complete a PhD without the support of others.
%As such, there are many people to thank for their help and contributions during my PhD, and I will do my best to thank them all here.
I am fortunate to have had an incredible community of people who have supported me throughout grad school. %and helped shape me into the researcher I am today. 

First, I must thank my advisor, Mark Oskin.
Mark has guided me through many different research directions, provided me with invaluable advice, and helped me grow into the researcher I am today.
Further thanks to Michael Taylor for his guidance and advice on the architecture side of things.
I am deeply grateful for their encouragement and support.

This dissertation presents work done by a large number of researchers that I have been lucky enough to collaborate with over the past six years.
Special thanks to Dustin Richmond and Max Ruttenberg for their help in developing the code generation framework at the heart of my dissertation work and for answering my many questions about hardware simulation.
Thank you to my collaborators on the \graphit side of things: Saman Amarasinghe, Julian Shun, Daniel Sanchez, Yunming Zhang, Ajay Brahmakshatriya, Victor Ying, Claire Hsu, and Changwan Hong.
Further thanks to Scott Davidson, Borna Ehsani, Dai Cheol Jung, Bandhav Veluri, and the rest of my collaborators on the \hb project. 

Many thanks to all of my labmates for not only their technical help and advice but also for making grad school fun; 
especially thanks to my friends Meghan Cowan, Ming Liu, Liang Luo, Amrita Mazumdar, Luis Vega, and Eddie Yang.
Thank you to my non-architect friends Kira Goldner, Lucy Lin, and Maarten Sap for the countless espresso room chats and weekend brunches.

Thank you to the entire Allen School staff for making the administrative and operational side of grad school run as smoothly as possible.
Many thanks to Elise Dorough especially for all of her support. 

Thank you to my undergraduate mentors and advisors: Michael Heroux, Imad Rahal, and Kris Nairn.
They introduced me to research and convinced me to pursue graduate studies.
Further thanks are due to Lynn Ziegler for encouraging me to take my first Computer Science course and forever changing the direction of my studies.

Thank you to my family: Kate, Terry, and John. 
Thank you for believing in me and encouraging me as I moved across the country to pursue a PhD.
To Beth, Joel, Hannah, and Julia, thank you for your support and for welcoming me so completely into your family.

Finally, immeasurable thanks to my wonderful partner, Andrew C. Schoen; 
I cannot imagine how different my graduate studies might have been without his encouragement and support.
\endgroup
