\chapter{Conclusion}\label{gen:sec:conclusion}
\markboth{Conclusion}{Conclusion}

This document presents proposed solutions to the challenge of writing peformant code for graph applications on manycore architectures. 
The optimizations presented in this work aim to better leverage the memory hierarchy and architectural features of a manycore in order to achieve high performance on graph applications. 
In this thesis proposal, I present a code generation backend to the \graphit DSL that allows for flexibility in application development and optimization. 
I show how existing optimizations can be implemented on a manycore in order to improve work efficiency and decrease workload imbalance across cores. 
I propose several manycore specific optimizations that target the memory hierarchy in order to more efficiently use memory bandwidth.
I further propose two areas of future work to increase the efficiency and performance of the code produced by my \graphit code generation backend. 

As the size of sparse graph data continues to grow, the importance of scalable, performant graph processing systems will only continue to increase. 
My thesis proposes ways that we can leverage the parallelism offered by emerging manycore architectures in order to maintain scalable performance on graph applications while also maintaining the flexibility and ease of programming offered through the use of the \graphit DSL. 
As graph processing demands grow and users turn more and more to emerging architectures, I expect that future work will continue to build on these efforts to reduce programming complexity through code generation and increase performance through optimizations that target architectural features. 

